Using HMMs for pre-snap coverage detection

### 2024-12-09 submission.Rmd
Amazon Prime Modell

Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take before each offensive play in American football. While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the tactics by the defence.
This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
By modeling hidden states that proxy the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.

>>>>

Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take before each offensive play in American football. While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the tactics by the defence.
This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
By modeling hidden states that represent the offensive player being guarded, this approach builds on previous attempts to predict zone- or man coverage (see screenshot below from the Pittsburgh Steelers vs. Cleveland Browns match), which focused primarily on specific plays without motion. In contrast, we are now able to include player movement and exploit the additional information available. In this way, we provide a data-driven framework for unravelling the complexity of defensive patterns, enabling real-time tactical insights for coaching and analysis.

<center>
  <img src="prime_coverage_prediction.jpg" width="600" height="400" alt="Coverage Prediction">
</center>
